Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-07 introduction.tex
Abstract and Introduction ready

At the beginning, there was darkness and then... bang! \ac{NC}
\cite{ahlswede2000network} appeared to save us from the evilness
of routing.

General introduction. Introduction to topic addressed in the journal.
Review of the State of the Art. Specify why our approach has benefits
and which are they. Indicate contributions.

Our work is organized as follows: Section \ref{sec:testbed} introduces
the testbed system, its setup, connectivity, files configuration and
compilation of the Kodo library. Section \ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section \ref{sec:metrics}
we describe the considered metrics for performance comparison of the
codes deployed in the \ac{Raspi}. In Section \ref{sec:measurements}, we show
the measurements in the \ac{Raspi} of the mentioned metrics. Final
conclusions and future work are reviewed in Section \ref{sec:conclusions}.

>>>>

%!TEX root = raspi_journal.tex
% At the beginning, there was darkness and then... bang! \ac{NC}
% \cite{ahlswede2000network} appeared to save us from the evilness
% of routing.
%
% General introduction. Introduction to topic addressed in the journal.
% Review of the State of the Art. Specify why our approach has benefits
% and which are they. Indicate contributions.

The emergence of powerful and inexpensive single-board computers
running full versions of a standard \ac{OS} from many possible in the last
years opens exciting possibilities for researchers. This not only
allows to seamlessly develop implementations that are compatible with
higher end devices and allows the reuse stable software from the Open
Source community. It also enables the deployment and testing of large
scale (hundreds of devices or more) distributed systems at a fraction
of the cost necessary in previous years. This paper focuses on
providing the critical steps and mechanisms followed to deploy, setup,
and maintain a large-scale testbed based on Raspberry Pi devices of
different models. We also complement this with an in-depth evaluation
of the Raspberry Pi as a platform to enable a new paradigm in
networking and storage known as \ac{NC}, which relies on
processing at intermediate nodes to achieve higher throughput, delay
and reliability performance than state-of-the-art routing techniques.

% INTRODUCTION TO NETWORK CODING
Introduced by Ahlswede et al.~\cite{ahlswede2000network}, network
coding constitutes a paradigm shift in the way how researchers and
industry understand and operate networks, by changing the role of
intermediate relays in the process of transmission of information.
Relays are no longer limited to storing and forwarding data, but also
take part in the coding process, through a process called recoding,
where the relay generates new linear combinations of incoming coded
packets without previously decoding the data. Network coding allows
the increase of throughput, reliability, security and delay
performance of the networks.

Compared to other broadly used coding schemes, e.g., Gallager's LDPC
codes~\cite{gallager1962low} or Reed-Solomon
codes~\cite{reed1960polynomial}, network coding is a technology that
has been implemented in real systems since the early years of its
conception. In 2006, Katabi et al.~\cite{katabi2006practical} published
the results of the performance of their protocol COPE when implemented
in a real wireless system, which relied on minimalistic coding to
provide interesting gains. In 2008, Pedersen et
al.~\cite{pedersen2008implementation} used commercially available
Symbian OS enabled mobile phones to implement network coding in a
\ac{D2D} cooperation-based application. In 2011, researchers at Aalborg
University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
coding library with open source code for researchers intended to make
network coding implementations easily available for both, the research
community and commercial entities.
Another case is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
IEEE 802.11 multi-hop meshed networks using some of the intuition from
COPE, but deploying it in real systems. It is currently available as
open source code in the Linux kernel. Many other implementations have
been tested on real world systems~
\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.

Despite this successful deployment in real systems, many of these
protocols and contributions have been implemented in separate testbeds
and the experiences are hard to reproduce.  Given the need for
reproducibility of past and future results among researchers and the
need of low-cost and easy-to-deploy testbeds, this paper contribution
is provide detailed instructions on how to set up a low-cost,
heterogeneous and scalable testbed consisting of Raspberry Pis
(Raspis), and by providing detailed measurements of goodput and energy
consumption of such testbed when performing network coding operations
with different codecs (such as full \ac{RLNC} \cite{ho2006random},
multicore-enabled \ac{RLNC}, sparse \ac{RLNC} and tunable sparse \ac{RLNC}).
The heterogeneity of the testbed comes from the fact that old, single core,
Raspberry Pi 1 model B units are integrated seamlessly in the described
testbed with newer, multicore versions, like the Raspberry Pi 2 model B.
Our testbed allows us to obtain a comprehensive set of measurements
to characterize the goodput and energetic footprint of these devices
showing that processing speeds of 800 Mbps and energy per bit values
of 1 nJ or even one order of magnitude less are possible.

Our work is organized as follows. Section~\ref{sec:testbed} introduces the
testbed system, its setup, files configuration, connectivity and compilation of
the Kodo library. Section~\ref{sec:schemes} defines the coding schemes employed
in our study. Later, in Section~\ref{sec:metrics} we describe the considered
metrics and methodology for performance comparison of the codes deployed in
the \ac{Raspi}. In Section~\ref{sec:measurements}, we show the measurements in
the \ac{Raspi} models of the mentioned metrics. Final conclusions and future work
are reviewed in Section~\ref{sec:conclusions}.

### 2016-06-07 raspi_journal.tex
Abstract and Introduction ready

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{Deploying a large-scale and configurable testbed for
  networking and storage can be challenging, not only due to the
  inherent costs of the hardware, but due to maintenance challenges
  and ability to replicate results consistently. The latter requires
  not only the devices to run the same operating system (OS), but also
  have exactly the same configurations and software packages. This
  paper presents the design, step-by-step setup, and configuration of
  an inexpensive testbed using Raspberry Pi devices for networking
  (wireless or wired) and storage applications, although the
  architecture can be used for other applications that require
  replicable results.  Specifically, we provide an in-depth
  performance evaluation for network coding, a new paradigm in
  networking and storage, and demonstrating the processing
  capabilities and energy behaviour of various code constructions and
  algorithms over different Raspberry Pi models. Our results show that
  processing speeds of more than 80 Mbps and 800 Mbps are attainable
  for a wide array of configurations for the Raspberry Pi Model 1 and
  2, respectively, and that processing energy per bit for network
  coding is below 1 nJ in a large regime and even an order of
  magnitude lower for some algorithms.}

>>>>

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{Deploying a large-scale and configurable testbed for
  networking and storage can be challenging, not only due to the
  inherent costs of the hardware, but due to maintenance challenges
  and ability to replicate results consistently. The latter requires
  not only the devices to run the same \ac{OS}, but also
  have exactly the same configurations and software packages. This
  paper presents the design, step-by-step setup, and configuration of
  an inexpensive testbed using Raspberry Pi devices for networking
  (wireless or wired) and storage applications, although the
  architecture can be used for other applications that require
  replicable results.  Specifically, we provide an in-depth
  performance evaluation for network coding, a new paradigm in
  networking and storage, and present the processing
  capabilities and energy behaviour of various code constructions and
  algorithms over different Raspberry Pi models. Our results show that
  processing speeds of more than 80 Mbps and 800 Mbps are attainable
  for a wide array of configurations for the Raspberry Pi Model 1 and
  2, respectively, and that processing energy per bit for network
  coding is below 1 nJ in a large regime and even an order of
  magnitude lower for some algorithms.}
